---
title: "Grade analysis"
author: "Arthur Lemoine"
format: html
editor: visual
execute:
  echo: false
  warning: false
  keep-md: true
---

```{r}
library(here)
here::i_am("grade101.Rproj")
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
library(vroom)
```

### Question 1

```{r}
grades <- vroom(here('grades.csv'))
```

### Question 2

```{r}
exam_summ <- grades |>
  summarise(
    Minimum = min(Exam, na.rm = TRUE),
    Maximum = max(Exam, na.rm = TRUE),
    Median = median(Exam, na.rm = TRUE),
    Mean = mean(Exam, na.rm = TRUE)
  )

exam_summ |> 
  knitr::kable(format = "markdown")
```

### Question 3

```{r}
st_no_exam <- sum(is.na(grades$Exam))
```

There are `r st_no_exam` students who did not take the exam.

### Question 4

```{r}
grades |> 
  ggplot(aes(x = Exam)) +
  geom_histogram(bins = 20, na.rm = TRUE) +
  labs(
    title = "Distribution of Exam Grades",
    x = "Exam Grade",
    y = "Frequency")
```

### Question 5

I decided not to remove NAs because they are still in a group (even though they didn't pass the exam).

```{r}
n_group <- grades |>
  group_by(Group) |>
  summarise(Count = n())

n_group|> 
  knitr::kable(format = "markdown")
```

### Question 6

```{r}
n_group |> ggplot(aes(x = Group, y = Count)) +
  geom_col() +
  labs(
    title = "Number of Students in Each Group",
    x = "Group",
    y = "Number of Students"
  ) + 
  theme(axis.text.x = element_text(angle = 90))
```

### Question 7

This one is clearly unreadble.

```{r}
grades |> ggplot(aes(x = Exam, fill = Group)) +
  geom_histogram(bins = 20, position = "dodge") +
  labs(
    title = "Distribution of Exam Grades by Group",
    x = "Exam Grade",
    y = "Frequency"
  )
```

```{r}
grades |> ggplot(aes(x = Group, y = Exam, fill = Group)) +
  geom_boxplot() +
  labs(
    title = "Box Plots of Exam Grades by Group",
    x = "Group",
    y = "Exam Grade") + 
  theme(axis.text.x = element_text(angle = 90))
```

### Question 8

```{r}
n_missed_exam <- grades |>
  group_by(Group) |>
  summarise("Missed exam count" = sum(is.na(Exam)))

n_missed_exam |> 
  knitr::kable(format = "markdown")
```

### Question 9

```{r}
n_missed_exam |> ggplot(aes(x = Group, y = `Missed exam count`, fill = Group)) +
  geom_col() +
  labs(
    title = "Number of Students Who Missed the Exam by Group",
    x = "Group",
    y = "Number of Students"
  ) + 
  theme(axis.text.x = element_text(angle = 90))
```

### Question 10

```{r}
grades_LF <- grades |>
  pivot_longer(
    cols = -c(Id, Group),
    names_to = "name",
    values_to = "value")
```

### Question 11

```{r}
n_miss_grades <- grades_LF |>
  group_by(Id) |>
  summarise(`Missing Grades` = sum(is.na(value)))
```

### Question 12

```{r}
n_miss_grades |> ggplot(aes(x = factor(`Missing Grades`))) +
  geom_bar() +
  labs(
    title = "Distribution of Missing Grades per Student",
    x = "Number of Missing Grades",
    y = "Count")
```

### Question 13

```{r}
n_missed_exam_LF <- grades_LF |>
  filter(name == "Exam" & is.na(value)) |>
  group_by(Group) |>
  summarise(`Missed exam count` = n())

n_missed_exam_LF |> 
  knitr::kable(format = "markdown")
```

### Question 14

```{r}
n_miss_online_test <- grades_LF |>
  filter(str_starts(name, "Online_MCQ_")) |>
  group_by(Id) |>
  summarise(`Missed online MCQ count` = sum(is.na(value)))
```

### Question 15

```{r}
n_miss_online_test_bis <- grades |>
  select(starts_with("Online_MCQ"), Id) |>
  rowwise() |>
  mutate(missed_MCQ = sum(is.na(c_across()))) |>
  select(Id, missed_MCQ)


grades_bis <- grades |>
  left_join(n_miss_online_test_bis, by = "Id")


ggplot(data = grades_bis, aes(x = as.factor(missed_MCQ), y = Exam, group = as.factor(missed_MCQ))) +
  geom_boxplot() +
  labs(
    title = "Distribution of Exam Grades conditioned by Missing Online Tests",
    x = "Number of Missing Online Tests",
    y = "Exam Grade") 
```

### Question 16

```{r}
missed_T_F <- grades |>
  select(Id, starts_with("MCQ_")) |>
  rowwise() |>
  mutate(Missed = any(is.na(c_across()))) |>
  select(Id, Missed)
```

### Question 17

```{r}
missed_pct <- grades |>
  select(Id, Group) |>
  left_join(missed_T_F, by = "Id") |>
  group_by(Group) |>
  summarise(Pct_missed = mean(Missed))
```

### Question 18
```{r}
avg_gr_grade <- grades |>
  group_by(Group) |>
  summarise(Avg_Exam_Grade = mean(Exam, na.rm = TRUE))

pct_avg_merged <- inner_join(missed_pct, avg_gr_grade, by = "Group")

ggplot(data = pct_avg_merged, aes(x = Pct_missed, y = Avg_Exam_Grade)) +
  geom_point() +
  labs(
    title = "Average Exam Grade per Group vs. Percentage of Missed MCQ Grade",
    x = "Percentage of Missed MCQ Grade",
    y = "Average Exam Grade")
```
We don't see any clear correlation between those two variables.
